Parallel Monte Carlo Synthetic Acceleration Methods for Discrete Transport Problems

### 2013-08-16 doc/dissertation/presentation/defense.tex
more cleanup in the defense slides

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Development of a linear solver for discrete systems leveraging
    Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Application to neutron transport
    \item Research is required to explore general solver development
    \item Performance is of concern
    \end{itemize}
    \bigskip
  \item Development of a nonlinear solver for discrete systems
    leveraging Monte Carlo Synthetic Acceleration
    \begin{itemize}
    \item Monte Carlo is a more natural fit
    \item Application to fluid flow
    \item Convergence of the linear model is of concern
    \end{itemize}
    \bigskip
  \item Parallelization of Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Scalability is of concern
    \end{itemize}
  \end{itemize}
\end{frame}

@@

@@

  \begin{equation}
    -\nabla \cdot \mathbb{D}_n \nabla \mathbb{U}_n + \sum_{m=1}^4
    \mathbb{A}_{nm} \mathbb{U}_m = \frac{1}{k} \sum_{m=1}^4
    \mathbb{F}_{nm} \mathbb{U}_n\ \ \ \ \ \ \ n = 1,2,3,4\:.
    \label{eq:spn_fission_matrix}
  \end{equation}

@@

  \begin{itemize}
  \item Swap in MCSA as the solver at each eigenvalue operation
    \medskip
  \item Use Thyra/Stratimikos interfaces for plug and play - Thanks
    Ross!
    \medskip
  \item Current Exnihilo implementation gives the full operator
  \end{itemize}

@@

      {\tiny
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{lccc}\hline\hline
            \multicolumn{1}{l}{Solver}&
            \multicolumn{1}{c}{1 Group}&
            \multicolumn{1}{c}{2 Groups}&
            \multicolumn{1}{c}{4 Groups}\\
            \hline
            BiCGStab-ILUT & 11.6 & 11.6 & 12.4 \\ 
            GMRES-ILUT & 18.1 & 17.9 & 18.9 \\
            MCSA-ILUT-R-C & 14.6 & 15.4 & 17.6 \\
            MCSA-ILUT-MR-C & 16.0 & 17.1 & 23.7 \\
            MCSA-ILUT-R-EV & 18.3 & 19.4 & 16.8 \\
            MCSA-ILUT-MR-EV & 19.6 & 22.4 & 17.5 \\
            Richardson-ILUT & 60.9 & 60.4 & 63.4 \\
            %%
            \hline\hline
          \end{tabular}
        \end{center}
        \caption{Average number of linear solver iterations per
          eigenvalue iteration.}
      \end{table}
      }

@@

      \begin{itemize}
        \item MCLS generates the same k-eigenvalue and neutron flux in
          all groups and spatial locations
          \bigskip
        \item MCSA converged in fewer iterations than GMRES, more
          iterations than BiCGStab
          \bigskip
        \item Explicit preconditioning strategy destroys sparsity and
          elevates CPU times (Ifpack ILUT)
          \bigskip
        \item Spectral radius and memory limitations combined to
          prevent solutions at finer discretizations
      \end{itemize}

@@

  \begin{itemize}
  \item Many physics problems of interest are nonlinear
    \bigskip
  \item Signficant research on Newton methods since the 1980's
    \bigskip
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Robust implementations
    \item No operator required
    \end{itemize}
    \bigskip
  \item Monte Carlo methods need the full operator
    \bigskip
  \item Automatic construction of the linear model is available
    \begin{itemize}
    \item Operator overloading for nonlinear residual differentiation
    \item Ideal for Monte Carlo
    \item Potential scaling improvements
    \item Resiliency benefits
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Drekar is a production physics suite that leverages Newton
    methods with FAD
  \item MCLS incorporated into the Drekar nonlinear solver scheme
    using Thyra/Stratimikos to implement FANM
  \item Thanks Roger!
  \end{itemize}

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      \multicolumn{1}{l}{Benchmark}& 
      \multicolumn{1}{c}{Newton-Krylov Speedup}\\
      \hline
      Convection, Ra=\sn{1}{3} & 338 \\
      Convection, Ra=\sn{1}{4} & 336 \\
      Convection, Ra=\sn{1}{5} & 346 \\
      Convection, Ra=\sn{1}{6} & 465 \\
      \hline
      Lid Driven, Re=100 & 299 \\
      Lid Driven, Re=300 & 322 \\
      Lid Driven, Re=500 & 288 \\
      Lid Driven, Re=700 & 488 \\
      \hline
      Backward Step, Re=200 & 400 \\
      Backward Step, Re=300 & 593 \\
      Backward Step, Re=400 & 825 \\
      Backward Step, Re=500 & 1057 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Newton-Krylov speedup over FANM.}  \textit{For all
      benchmarks the explicit MCSA preconditioning strategy caused
      significantly larger CPU times for FANM when compared to the
      Newton-Krylov solutions.}}
  \label{tab:benchmark_time_comparison}
\end{table}

>>>>

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Development of a linear scheme for discrete systems leveraging
    Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Application to neutron transport
    \item Research is required to explore general MCSA development
    \item Performance is of concern
    \end{itemize}
    \bigskip
  \item Development of a nonlinear scheme for discrete systems
    leveraging Monte Carlo Synthetic Acceleration
    \begin{itemize}
    \item Monte Carlo is a more natural fit
    \item Application to fluid flow
    \item Convergence of the linear model is of concern
    \end{itemize}
    \bigskip
  \item Parallelization of Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Scalability is of concern
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \smallskip

@@

  \smallskip

  \begin{equation}
    -\nabla \cdot \mathbb{D}_n \nabla \mathbb{U}_n + \sum_{m=1}^4
    \mathbb{A}_{nm} \mathbb{U}_m = \frac{1}{k} \sum_{m=1}^4
    \mathbb{F}_{nm} \mathbb{U}_m\ \ \ \ \ \ \ n = 1,2,3,4\:.
    \label{eq:spn_fission_matrix}
  \end{equation}

@@

  \begin{itemize}
  \item Swap in MCSA as the solver at each eigenvalue operation
    \medskip
  \item Use Thyra/Stratimikos interfaces for plug and play with $SP_N$
    \medskip
  \item Current Exnihilo implementation gives the full operator
  \end{itemize}

@@

      {\tiny
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{lccc}\hline\hline
            \multicolumn{1}{l}{Solver}&
            \multicolumn{1}{c}{1 Group}&
            \multicolumn{1}{c}{2 Groups}&
            \multicolumn{1}{c}{4 Groups}\\
            \hline
            BiCGStab-ILUT & 11.6 & 11.6 & 12.4 \\ 
            GMRES-ILUT & 18.1 & 17.9 & 18.9 \\
            MCSA-ILUT-R-C & 14.6 & 15.4 & 17.6 \\
            MCSA-ILUT-MR-C & 16.0 & 17.1 & 23.7 \\
            MCSA-ILUT-R-EV & 18.3 & 19.4 & 16.8 \\
            MCSA-ILUT-MR-EV & 19.6 & 22.4 & 17.5 \\
            Richardson-ILUT & 60.9 & 60.4 & 63.4 \\
            %%
            \hline\hline
          \end{tabular}
        \end{center}
        \caption{ Average number of linear solver iterations per
          eigenvalue iteration.}
      \end{table}
      }

@@

      \begin{itemize}
        \item MCLS generates the same k-eigenvalue and neutron flux in
          all groups and spatial locations
          \bigskip
        \item MCSA converged in fewer iterations than GMRES, more
          iterations than BiCGStab
          \bigskip
        \item Explicit preconditioning strategy destroys sparsity and
          elevates CPU times (Ifpack ILUT)
          \bigskip
        \item Spectral radius and memory limitations combine to
          prevent solutions at finer discretizations
      \end{itemize}

@@

  \begin{itemize}
  \item Many physics problems of interest are nonlinear
    \bigskip
  \item Signficant research on Newton methods since the 1980's
    \bigskip
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Robust implementations
    \item No operator required
    \end{itemize}
    \bigskip
  \item Monte Carlo methods need the full operator
    \bigskip
  \item Automatic construction of the linear model is available
    \begin{itemize}
    \item Operator overloading for nonlinear residual differentiation
    \item Ideal for Monte Carlo
    \item Potential scaling improvements
    \item Potential resiliency benefits
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Drekar is a production physics suite that leverages Newton
    methods with FAD
  \bigskip
  \item MCLS incorporated into the Drekar nonlinear solver scheme
    using Thyra/Stratimikos to implement FANM
  \end{itemize}

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      \multicolumn{1}{l}{Benchmark}& 
      \multicolumn{1}{c}{NK Speedup}\\
      \hline
      Convection, Ra=\sn{1}{3} & 338 \\
      Convection, Ra=\sn{1}{4} & 336 \\
      Convection, Ra=\sn{1}{5} & 346 \\
      Convection, Ra=\sn{1}{6} & 465 \\
      \hline
      Lid Driven, Re=100 & 299 \\
      Lid Driven, Re=300 & 322 \\
      Lid Driven, Re=500 & 288 \\
      Lid Driven, Re=700 & 488 \\
      \hline
      Backward Step, Re=200 & 400 \\
      Backward Step, Re=300 & 593 \\
      Backward Step, Re=400 & 825 \\
      Backward Step, Re=500 & 1057 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Newton-Krylov speedup over FANM.}  \textit{For all
      benchmarks the explicit MCSA preconditioning strategy caused
      significantly larger CPU times for FANM when compared to the
      Newton-Krylov solutions.}}
  \label{tab:benchmark_time_comparison}
\end{table}
